---
title: "covid_project"
author: "Jaehee Lee"
date: "11/27/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Task1: What days did states have their first case and the rate?
Graph the rates in a plot with respect to time. Try different kinds of plots potentially coloring or categorizing by a variable suca as state or region. 
```{r}
library(COVID19)
library(tidyverse)
library(readr)
library(lubridate)
```

```{r}
covid19("US", level = 2) %>%
  select(-c("administrative_area_level_1","administrative_area_level_3","latitude","longitude","currency", "administrative_area_level_1")) %>% 
  relocate("administrative_area_level_2", "date", "confirmed", "tests")-> easy_covid_df
```


```{r}
easy_covid_df %>% 
  group_by(administrative_area_level_2, confirmed) %>% 
  # mutate(first_date = min(date)) %>% 
  select(administrative_area_level_2, date, confirmed) %>% 
  arrange(date, administrative_area_level_2) %>% 
  group_by(administrative_area_level_2) %>% 
  filter(confirmed>0) %>% 
  mutate(daily_rate = c(0,diff(confirmed))) %>% 
  mutate(second_rate= c(0,diff(daily_rate))) %>% 
  arrange(administrative_area_level_2, date) -> maindata
```   

# Get mean of confirmed cases for each States
```{r}
library(data.table)
meandata <- setDT(maindata)[,list(Mean = as.numeric(mean(confirmed))), by=administrative_area_level_2]
```

## inner join with maindata and meandata
```{r}
inner_join(meandata, maindata, by = "administrative_area_level_2") -> a
```

## Plot mean for every states
```{r}
ggplot(data = a, aes(x = Mean, y = administrative_area_level_2))+
  geom_point()+
  theme_bw()
```

# What day did states have their first case? 
```{r}
library(dplyr)
maindata %>%
    group_by(administrative_area_level_2) %>%
    arrange(date) %>%
    slice(1L) %>%  
    select(administrative_area_level_2, date) -> first_case_day

```
# show the states n
```{r}
unique(maindata$administrative_area_level_2)
```
# Northeast
```{r}
# Northeast 
# 11 states there are 
#Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont, Delaware, Maryland, New #Jersey, New York, and Pennsylvania.
maindata %>%  
  filter(administrative_area_level_2 == c("Connecticut", "Maine", "Massachusetts", "New Hampshire","Rhode Island", "Vermont", "Delaware", "Maryland", "New Jersey", "New York", "Pennsylvania")) -> northeast
```
```{r}
library(dplyr)
northeast %>%
    group_by(administrative_area_level_2) %>%
    arrange(date) %>%
    slice(1L) %>%  
    select(administrative_area_level_2, date) -> northeast_first_case

# glimpse(northeast_first_case$date)
northeast_first_date <- as.Date(northeast_first_case$date)
# yday is the function to check how many days past after this year. 
northeast_convert_days <- yday(northeast_first_date) - 1 # January 1 = day 0 
mean_of_northeast <- mean(northeast_convert_days)
mean_NE_start_date <- as.Date(mean_of_northeast, origin = "2020-01-01")


```
# Northeast first case date mean 
```{r}
mean_NE_start_date
```

```{r}
ggplot(data =northeast, mapping = aes(x = date , y = confirmed))+ 
  stat_smooth(method = loess,color='black') +
  labs(x = "Time", y= "Confirmed Cases")+
   geom_vline(xintercept = mean_NE_start_date, color = "red", linetype = 2)+
  theme_bw()  -> northeast_confirmed

ggplot(data =northeast, mapping = aes(x = date , y = daily_rate))+ 
  stat_smooth(method = loess,color='black') +
  labs(x = "Time", y = "Daily Rate")+
     geom_vline(xintercept = mean_NE_start_date, color = "red", linetype = 2)+
  theme_bw()  -> northeast_rate

ggplot(data =northeast, mapping = aes(x = date , y = second_rate))+ 
  stat_smooth(method = loess,color='black') +
  labs(x = "Time", y = "Second Rate")+
       geom_vline(xintercept = mean_NE_start_date, color = "red", linetype = 2)+
  theme_bw()  -> northeast_second_rate


```

# Midwest
```{r}
# 12 states 
maindata %>%  
  filter(administrative_area_level_2== c("Ohio","Michigan","Indiana","Wisconsin","Illinois","Minnesota","Iowa","Missouri","North Dakota", "South Dakota","Nebraska","Kansas")) -> midwest
  
```



```{r}
library(dplyr)
midwest %>%
    group_by(administrative_area_level_2) %>%
    arrange(date) %>%
    slice(1L) %>%  
    select(administrative_area_level_2, date) -> midwest_first_case



# glimpse(midwest_first_case$date)
midwest_first_date <- as.Date(midwest_first_case$date)
# yday is the function to check how many days past after this year. 
midwest_convert_days <- yday(midwest_first_date) - 1 # January 1 = day 0 
mean_of_midwest <- mean(midwest_convert_days)
mean_MW_start_date <- as.Date(mean_of_midwest, origin = "2020-01-01")
```

# Midwest first case date mean
```{r}
mean_MW_start_date
```


```{r}
ggplot(data =midwest, mapping = aes(x = date , y = confirmed))+ 
  stat_smooth(method = loess,color='dark blue') +
  labs(x = "Time", y= "Confirmed Cases")+
   geom_vline(xintercept = mean_MW_start_date, color = "red", linetype = 2)+
  theme_bw() -> midwest_confirmed

ggplot(data =midwest, mapping = aes(x = date , y = daily_rate))+ 
  stat_smooth(method = loess,color='dark blue') +
   labs(x = "Time", y = "Daily Rate")+
   geom_vline(xintercept = mean_MW_start_date, color = "red", linetype = 2)+
  theme_bw() -> midwest_rate

ggplot(data =midwest, mapping = aes(x = date , y = second_rate))+ 
  stat_smooth(method = loess,color='dark blue') +
   labs(x = "Time", y = "Second Rate")+
   geom_vline(xintercept = mean_MW_start_date, color = "red", linetype = 2)+
  theme_bw() -> midwest_second_rate


```


# South
```{r}
#16states
maindata %>%  
  filter(administrative_area_level_2==c("Delaware", "Maryland", "Virginia", "West Virginia", "Kentucky", "North Carolina", "South Carolina", "Tennessee", "Georgia", "Florida", "Alabama", "Mississippi", "Arkansas", "Louisiana", "Texas", "Oklahoma")) -> south

```
```{r}
library(dplyr)
south %>%
    group_by(administrative_area_level_2) %>%
    arrange(date) %>%
    slice(1L) %>%  
    select(administrative_area_level_2, date) -> south_first_case

# glimpse(south_first_case$date)
south_first_date <- as.Date(south_first_case$date)
# yday is the function to check how many days past after this year. 
south_convert_days <- yday(south_first_date) - 1 # January 1 = day 0 
mean_of_south <- mean(south_convert_days)
mean_S_start_date <- as.Date(mean_of_south, origin = "2020-01-01")
```
# South first case day mean
```{r}
mean_S_start_date
```



```{r}
ggplot(data =south, mapping = aes(x = date , y = confirmed))+ 
  stat_smooth(method = loess,color='dark green') +
  labs(x = "Time", y= "Confirmed Cases")+
  geom_vline(xintercept = mean_S_start_date, color = "red", linetype = 2)+
  theme_bw() -> south_confirmed

ggplot(data =south, mapping = aes(x = date , y = daily_rate))+ 
  stat_smooth(method = loess,color='dark green') +
 # geom_vline(xintercept = mdy("06/03/20"))+
   labs(x = "Time", y = "Daily Rate")+
    geom_vline(xintercept = mean_S_start_date, color = "red", linetype = 2)+
  theme_bw() -> south_rate



ggplot(data =south, mapping = aes(x = date , y = second_rate))+ 
  stat_smooth(method = loess,color='dark green') +
   labs(x = "Time", y = "Second Rate")+
    geom_vline(xintercept = mean_S_start_date, color = "red", linetype = 2)+
  theme_bw() -> south_second_rate


```



# West
```{r}
#13states
maindata %>%  
  filter(administrative_area_level_2 == c("Montana", "Idaho", "Wyoming", "Colorado", "New Mexico", "Arizona", "Utah", "Nevada", "California", "Oregon", "Washington", "Alaska", "Hawaii")) -> west

```
```{r}
library(dplyr)
west %>%
    group_by(administrative_area_level_2) %>%
    arrange(date) %>%
    slice(1L) %>%  
    select(administrative_area_level_2, date) -> west_first_case



# glimpse(west_first_case$date)
south_first_date <- as.Date(west_first_case$date)
# yday is the function to check how many days past after this year. 
west_convert_days <- yday(south_first_date) - 1 # January 1 = day 0 
mean_of_west <- mean(west_convert_days)
mean_W_start_date <- as.Date(mean_of_west, origin = "2020-01-01")

```

# West first case date mean 
```{r}
mean_W_start_date
```



```{r}
ggplot(data =west, mapping = aes(x = date , y = confirmed)) + 
  stat_smooth(method = loess,color='orange') +
  labs(x = "Time", y= "Confirmed Cases")+
      geom_vline(xintercept = mean_W_start_date, color = "red", linetype = 2)+
  theme_bw() -> west_confirmed 

ggplot(data =west, mapping = aes(x = date , y = daily_rate)) + 
  stat_smooth(method = loess,color='orange') +
   labs(x = "Time", y = "Daily Rate")+
      geom_vline(xintercept = mean_W_start_date, color = "red", linetype = 2)+
  theme_bw() -> west_rate 

ggplot(data =west, mapping = aes(x = date , y = second_rate)) + 
  stat_smooth(method = loess,color='orange') +
   labs(x = "Time", y = "Second Rate")+
      geom_vline(xintercept = mean_W_start_date, color = "red", linetype = 2)+
  theme_bw() -> west_second_rate

```


# Four Regions 
# Confirmed vs Time
```{r}
library(ggplot2)
library(ggpubr)
```


```{r}
regions_confirmed <- ggarrange(northeast_confirmed,midwest_confirmed,south_confirmed, west_confirmed,
                    labels = c("Northeast", "Midwest", "South", "West"),
                    ncol = 2, nrow = 2)
regions_confirmed
```
# Four Regions 
# Daily_rate vs Time 
```{r}
regions_rate <- ggarrange(northeast_rate, midwest_rate,south_rate, west_rate,
                    labels = c("Northeast", "Midwest", "South", "West"),
                    ncol = 2, nrow = 2)
regions_rate
```
# Four Regions 
# Second_rate vs Time 
```{r}
regions_second_rate <- ggarrange(northeast_second_rate,midwest_second_rate,south_second_rate, west_second_rate,
                    labels = c("Northeast", "Midwest", "South", "West"),
                    ncol = 2, nrow = 2)
regions_second_rate
```

