---
title: "Task 4"
output: html_notebook
---


Task 4: Compare how quickly states acted and how their covid rates are X number of weeks after.  Get a proportion based on the amount of time. 
Meaning day 1 is the first day of a covid case in the us. Day n meaning last day of the count is today's date november 26. This difference will give a positive number most likely around 250days call .
4a. Find the difference for the first date that covid arrived in us and todays date call this variable daymax it should be the same for each state.
4b.For each state find the difference between  their first stay home restriction and first day of covid in us. 
4c. Get the proportion of day max dif for each state.
```{r}
library(tidyverse)
library(lubridate)
library(COVID19)
```


```{r}
covid19("US", level = 2) %>%
  rename(state_name = administrative_area_level_2 ) %>% 
  select(-c("administrative_area_level_1","administrative_area_level_3","latitude","longitude","currency", "administrative_area_level_1")) %>% 
  relocate("state_name", "date", "confirmed", "tests")-> easy_covid_df
```

https://www.nejm.org/doi/full/10.1056/NEJMoa2001191
first day: January 19, 2020
```{r}
mdy("January 19, 2020")->day1
today()-day1 ->tot_days0

easy_covid_df %>% 
  filter(!)
  group_by(state_name, confirmed) %>% 
  mutate(first_case = min(confirmed)) %>% 
  select(state_name, date, first_case, confirmed) %>% 
  arrange(date, state_name)


  
```


Here we have the first day of state_home restrictions for for each state.
```{r}
easy_covid_df %>% 
  group_by(state_name, stay_home_restrictions) %>% 
  mutate(first_date0 = min(date)) %>% 
  select(state_name, date, first_date0, stay_home_restrictions) %>% 
  distinct(first_date0) %>% 
  filter(!is.na(stay_home_restrictions)&(!stay_home_restrictions==0)) %>% 
  arrange(first_date0) %>% 
  group_by(first_date0) %>% 
  mutate(num_of_states_with_with_rest=n()) %>% 
  arrange(state_name) %>% 
  group_by(state_name) %>% 
  filter( first_date0 ==min(first_date0))->first_date_stayh_orders
  


```



